---
title: "433hw3"
author: "Yifan Zhang"
date: "2/28/2021"
output: github_document
---

```{r}
library(dplyr)
library(nycflights13)
library(ggplot2)

flights
?flights
flights[,5:10]
colnames(flights)
str(flights)
```

## Problem

##### In r4ds flights… What time of day should you fly if you want to avoid delays as much as possible? 

```{r}
data <- flights #load a data frame for observing the data

flights %>%
  group_by(hour) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = T), cancel_ratio = sum(is.na(dep_delay))/length(dep_delay))
```
From this result, we see that we should better fly from 5:00 a.m. to 12:00 p.m. During this period, the average delay time is less than 9 minutes, and the ratio of cancelled flights is less than 2%.

##### Does this choice depend on anything? Season? Weather? Airport? Airline? Find three patterns (“null results” are ok!). Write your results into Rmarkdown. Include a short introduction that summarizes the three results. Then, have a section for each finding. Support each finding with data summaries and visualizations. Include your code when necessary.

### Introduction

Given that we should better fly from 5:00 a.m. to 12:00 p.m. to avoid delays as much as possible, in the later work, I will explore how this choice is related to other factors, including airlines, departed airports, and seasons. From the below analysis, we know that all of these three factors will affect the average delay time of the flights. However, airlines and seasons have a greater impact on choosing flights during the given period than departed airports.

### Airlines

In this section, I will try to explore how this choice depends on different airlines.

```{r}
#data frame
f1 <- flights %>%
  filter(hour >= 5 & hour <= 11) %>%
  group_by(carrier) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = T)) %>%
  arrange(avg_delay)
f1

# visualization
ggplot(f1, aes(x = carrier, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "#2e2e97") +
  labs(
    x = "Carrier",
    y = "Average Delay Time"
  )
```

From the data summaries and visualization above, we see airlines are important while choosing a flight during the given period. AS(the best choice), US, AA, DL, and B6 are better choice this time.

### Departed Airports

In this section, I will try to explore how this choice depends on different departed airports (origins). 

```{r}
#data frame
f2 <- flights %>%
  filter(hour >= 5 & hour <= 11) %>%
  group_by(origin) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = T)) %>%
  arrange(avg_delay)
f2

# visualization
ggplot(f2, aes(x = origin, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "#972e30") +
  labs(
    x = "Departed Airport",
    y = "Average Delay Time"
  )
```

From the data summaries and visualization above, we see average delay time of different airports also differs, though the difference is less than 3 minutes. But the LGA is still the place with the lowest average delay times, and it is followed by JFK and EWR. 

### Seasons

In this section, I will try to explore how this choice depends on seasons. 

```{r}
#data frame
f3 <- flights %>%
  filter(hour >= 5 & hour <= 11) %>%
  mutate(season = ifelse(month >= 3 & month <= 5, 'Spring',
                         ifelse(month >= 6 & month <= 8, 'Summer',
                                ifelse(month >= 9 & month <= 11, 'Autumn', 'Winter')))) %>% #add a new column called 'season'
  group_by(season) %>%
  summarise(avg_delay = mean(dep_delay, na.rm = T)) %>%
  arrange(avg_delay)
f3

# visualization
ggplot(f3, aes(x = season, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "#C8e356") +
  labs(
    x = "Season",
    y = "Average Delay Time"
  ) 
```

From the data summaries and visualization above, we see seasons are really important while choosing a flight from 5:00 a.m. to 12:00 p.m..In autumn, the average delay time is really short, which is less than 1 minute. But in winter, the average delay time is much more, which is over 6 minutes. Therefore, we better fly in autumn(Sep. to Nov.).
